### 2020-01-28 presentations/spsa.Rmd

---
title: "Felony Disenfranchisement and Neighborhood Turnout"
subtitle: "The Case of New York City"
shorttitle: "Other"
author: Kevin Morris
abstract: kevin.morris@nyu.edu
institute: Brennan Center for Justice
date: Southern Political Science Association, 2020
output:
  beamer_presentation:
    latex_engine: xelatex
    theme: "Berlin"
    keep_tex: true
bibliography: "../markdown/citations.bib"
---
```{r setup, echo = F, message = F, warning = F}
knitr::opts_knit$set(dev = "cairo_pdf")
library(systemfit)
library(AER)
library(sandwich)
library(lmtest)
library(scales)
library(tidyverse)
library(extrafont)
library(kableExtra)
library(stargazer)
library(data.table)
loadfonts()
```

## Outline

>* Introducing space into the conversation

>* Redefining "lost voters"

>* Identifying neighborhoods with lost voters

>* Testing turnout effects

## Introducing Space

>* Most of the existing literature on felony disenfranchisement and indirect turnout effects looks for effects at the state level [@Miles2004; @Bowers2009; @King2016]

>* But we *know* that incarceration patterns aren't uniformly distributed throughout a given state [see, for instance, @Gelman2007]

>* Turnout effects are likely to be socially mediated - and therefore show up in the neighborhoods where disenfranchised individuals live [e.g. @Foladare1968; @Huckfeldt1979; @Cho2006]

## Redefining Lost Voters

>* Much of the existing literature cannot identify individuals who *would have voted* if not for their disenfranchisement [@Burch2013]

>* This undermines our ability to distinguish the causal effects of incarceration / probation from the effects of disenfranchisement

>* This is especially likely given the low turnout propensity among the formerly disenfranchised [e.g. @Meredith2015; @Gerber2017; @White2019]

## This Project Addresses Both Problems

>* Administrative data from the NYS DOC allows us to conduct individual-level analysis

>* Purge records in the NYS voter file allow us to construct vote histories for disenfranchised individuals

>* Lost voters are all formally disenfranchised individuals who have voted in the past 10 years

## Identifying Neighborhoods with Lost Voters

```{r ttt, echo=FALSE, out.width = '100%'}
knitr::include_graphics("../temp/lost_voters_map.png")
```

## Methodology

* A genetic match algorithm [@Sekhon2011] is used to match treated census block groups to untreated ones.

* Each block group is matched to 30 untreated block groups; matches are done with replacement.

## Matching Results

```{r, echo=FALSE,out.width="49%", out.height="49%",fig.show='hold',fig.align='center'}
knitr::include_graphics(c("../temp/perc_white_spsa.png", "../temp/perc_black_spsa.png"))
```

## Matching Results

```{r, echo=FALSE,out.width="49%", out.height="49%",fig.show='hold',fig.align='center'}
knitr::include_graphics(c("../temp/perc_latino_spsa.png", "../temp/income_spsa.png"))
```

## Testing Turnout Effects

```{r, echo=FALSE,out.width="85%", out.height="85%",fig.show='hold',fig.align='center'}
knitr::include_graphics(c("../temp/coef_plot1.png"))
```
## Testing Turnout Effects

```{r, echo=FALSE,out.width="85%", out.height="85%",fig.show='hold',fig.align='center'}
knitr::include_graphics(c("../temp/coef_plot2.png"))
```

## Testing Turnout Effects

```{r, echo=FALSE,out.width="85%", out.height="85%",fig.show='hold',fig.align='center'}
knitr::include_graphics(c("../temp/coef_plot3b.png"))
```

## Testing Turnout Effects

```{r post-est, echo=FALSE, out.width = '100%'}
knitr::include_graphics(c("../output/dep_map.png"))
```

## Conclusions

>* The indirect turnout effects identified in past research appear to be geographically concentrated where the lost voters live.

>* @Hajnal2009 and others have demonstrated that turnout differentials can have real political implications, especially in low-turnout contests.

>* Felony disenfranchisement undermines Black political representation at the local level.

## We Made It!

Thanks!

kevin.morris@nyu.edu

## References {.allowframebreaks}