### 2020-07-06 markdown/part2/registration_regs.Rmd

---
output:
  bookdown::pdf_document2:
    toc: yes
    fig_caption: yes
    latex_engine: xelatex
    keep_tex: true
bibliography: "../welcome_home_vote.bib"
link-citations: yes
fontsize: 12pt
header-includes:
    - \usepackage{rotating}
    - \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{A\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{A\arabic{figure}}}
    - \usepackage{setspace}
---
\newpage
\doublespace
```{r setup, echo = F, message = F, warning = F}
knitr::opts_knit$set(root.dir = "../..", dev = "cairo_pdf")
library(systemfit)
library(AER)
library(sandwich)
library(lmtest)
library(scales)
library(tidyverse)
library(extrafont)
library(kableExtra)
library(data.table)
library(stargazer)
loadfonts()
```
# Treatment Effect on Registration {-}

In the body of this manuscript, I argue that restoring voting rights to individuals in an in-person meeting while they are still on parole increases their eventual propensity to vote. The mechanism for this effect, of course, must run through voter registration; parole officers are encouraging their stewards to register, and the actual decision to vote occurred when the individual was no longer on parole. Voter registration is, in and of itself, relatively uninteresting in this context: if the executive order had been successful at registering formerly incarcerated individuals, but did not result in higher turnout, it can hardly be said to have been effective at increasing political representation.

In this Supplemental Appendix, I re-estimate the primary analysis found in Table 4 and 5in the main body of this manuscript. Here, however, I use registration as the dependent variable, rather than turnout in 2018.

\begin{singlespace}
```{r race-reg-chunk, echo = F, results = "hide", message = F, warning = F}
load("./temp/app_reg.rdata")

stargazer(model1, model2,
          header = F,
          type = "text", notes.align = "l",
          covariate.labels = c("D(Discharged After EO 181)", "D(Male)", "Age (Years)",
                               "Voted in 2008",
                               "Years on Parole",
                               "Days off Parole Before Election",
                               "Black $\\times$ D(Discharged After EO 181)"),
          dep.var.labels = c("Registered to Vote"),
          title = "\\label{tab:racevar} ITT By Race",
          table.placement = "H",
          omit = c("race", "felony"),
          keep.stat = c("n", "rsq", "adj.rsq"),
          add.lines=list(c("Race / Ethnicity FE", "", "X", "X", "X"),
                         c("Felony Class FE", "", "", "X", "X")),
          table.layout = "-dm#-t-a-s-n",
          out = "./temp/test.tex",
          notes = "TO REPLACE",
          out.header = F,
          se = list(m1_ses$V1, m2_ses$V1))

j <- fread("./temp/test.tex", header = F, sep = "+")

note.latex <- "\\multicolumn{2}{l}{\\scriptsize{\\parbox{.5\\linewidth}{\\vspace{2pt}$^{***}p<0.01$, $^{**}p<0.05$, $^*p<0.1$. \\\\Robust standard errors in parentheses.}}}"

j <- j %>% 
  mutate(V1 = ifelse(grepl("TO REPLACE", V1), note.latex, V1),
         V1 = ifelse(grepl("\\\\#tab", V1), gsub("\\\\#", "", V1), V1)) %>% 
  filter(!grepl("Note:", V1)) %>% 
  mutate(n = row_number())

insert1 <- "\\fontsize{10}{12}\\selectfont"

j <- bind_rows(j, data.frame(V1 = c(insert1), n = c(1.1))) %>% 
  arrange(n) %>% 
  select(-n)

write.table(j, "./temp/tablexxx.tex", quote = F, col.names = F,
            row.names = F)

```
\input{"../../temp/tablexxx.tex"}
\end{singlespace}

As with turnout, there is a positive ITT of around 1 percentage point. However, as Model 2 makes clear, this was not so for Black voters, whose registration was apparently undisturbed by the treatment.